---
title: "StepReg Vignette"
author:
- name: Junhui Li
  affiliation: CAU and Umass
  email: junhuili@cau.edu.cn
package: StepReg
output:
  BiocStyle::html_document
abstract: |
  Stepwise regression is a popular data-mining method to identify a useful subset of the predictors to be used in a multiple regression model. We developed an R package StepReg, which allow users to perform linear, logistic and cox proportional hazard stepwise regression with a widely used selection criteria and stop rules are available in forward selection, backward elimination, both-direcition and best subset method. User can specify effects to be included in all models and do multivariate multiple stepwise regression.
vignette: |
  %\VignetteIndexEntry{StepReg Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

Here we applied StepReg to the well-known mtcars data and lung data for clarifying how to perform linear, logistic and Cox stepwise regression.

```{r library}
#install.package("StepReg")
library(StepReg)
```

## linear stepwise regression with data mtcar

* 1) linear stepwise regression using 'forward' method for variable selection and 'AIC' as criteria for stop rules
```{r}
data(mtcars)
formula <- mpg ~ .
sForwAIC <- stepwise(formula=formula,
         data=mtcars,
         selection="forward",
         select="AIC")
sForwAIC
```

From the above result, we can see that stepwise() output a list with 5 tables. 



```{r sessionInfo, echo=FALSE}
sessionInfo()
```
